---
title: "2022 W 1 preppin' data Challenge - R Solution"
author: "Paula Munoz"
date: "2022/01/17"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2022: Week 2 -The Prep School - Birthday Cakes

Challenge by Carl Allchin.

Challenge Link:
<https://preppindata.blogspot.com/2022/01/2022-week-2-prep-school-birthday-cakes.html>

Data Source:
<https://drive.google.com/file/d/1p8gt3cR3ATCeGK81pnT90x0a6dbCXst1/view>

### DETAILS

#### Requirements

-   Input the csv file (link above)
-   Remove any unnecessary fields 
-   Form the pupil’s name in <First Name> <Last Name> format
-   Create the date for the pupil's birthday in calendar year 2022
-   Work out what day of the week the pupil's birthday falls on, Remember if the birthday falls on a Saturday or Sunday, we need to change the weekday to Friday
-   Work out what month the pupil's birthday falls within
-   Count how many birthdays there are on each weekday in each month
-   Remove any unnecessary columns of data
-   Output the data


## LOADING LIBRARIES 

```{r pressure, message=FALSE}
library(tidyverse)
library(skimr)
library(knitr)
library(tidyquant)
library(lubridate)
```


## IMPORTING DATA 

Data resides in a .csv file

```{r}
input <- readr::read_csv('additional_files/PD 2022 Wk 1 Input - Input.csv')

```


## INSPECTING DATA 

* Inspecting first five rows:
```{r}
input %>%
  head() %>%
  kable()
```


* Detailed summary:
```{r}
input %>%
  skim()
```


* Glimpse data:
```{r}
input %>%
  glimpse()
```


## DATA WRANGLING

### Remove unnecessary columns
```{r}
tidy_data  <- input %>% select(-`Parental Contact Name_1`, -`Parental Contact Name_2`, -`Preferred Contact Employer`, -`Parental Contact`, -gender)

tidy_data%>%
  head(10) %>%
  kable() 
```



### Form the pupil’s name 

```{r}

tidy_data <- tidy_data %>%
  mutate(Pupils_Name = str_c(`pupil first name`, ', ',`pupil last name`))

tidy_data%>%
  head(10) %>%
  kable() 
```
```{r}
tidy_data %>% glimpse()
```


### 2. Create the date for the pupil's birthday in calendar year 2022 (not academic year) 

```{r}

tidy_data <- tidy_data %>%
  # Convert Date of Birth to Date Format
  mutate(date = (mdy(tidy_data$`Date of Birth`))) %>%
  # Extract year date
  mutate(year_date = year(date)) %>%
  # Mutate "This Year's Birthday"
  mutate(This_years_birthday = case_when(
                                  year_date == 2011 ~ date + years(11),
                                  year_date == 2012 ~ date + years(10),
                                  year_date == 2013 ~ date + years(9),
                                  year_date == 2014 ~ date + years(8),
                                  year_date == 2015 ~ date + years(7)
  ))
  
tidy_data%>%
  head(10) %>%
  kable() 
```

### 3. Work out what day of the week the pupil's birthday falls on. Remember if the birthday falls on a Saturday or Sunday, we need to change the weekday to Friday

```{r}

tidy_data <- tidy_data %>%
  mutate(cake_needed_on = case_when(
    wday(This_years_birthday, label = TRUE, abbr = FALSE) == "Saturday" ~ "Friday",
    wday(This_years_birthday, label = TRUE, abbr = FALSE) == "Sunday" ~ "Friday",
    TRUE ~ as.character(wday(This_years_birthday, label = TRUE, abbr = FALSE)))
    )
           
tidy_data%>%
  head(10) %>%
  kable()            

```

### 4. Work out what month the pupil's birthday falls within

```{r}
tidy_data <- tidy_data %>%
  mutate(Month = month(date, label = TRUE , abbr = FALSE))

tidy_data%>%
  head(10) %>%
  kable() 
```
### 5. Count how many birthdays there are on each weekday in each month 

```{r}
  tidy_data <- tidy_data %>%
  group_by(Month, cake_needed_on) %>%
  summarise(BDs_per_weekday_and_month = n()) %>%
  ungroup() %>%
  right_join(tidy_data) 

tidy_data%>%
  head(10) %>%
  kable() 
```

### 6. Remove unnecessary columns

```{r}
tidy_data <- tidy_data %>%
  select(Pupils_Name, `Date of Birth`, This_years_birthday, Month, cake_needed_on, BDs_per_weekday_and_month)

tidy_data%>%
  head(10) %>%
  kable() 
```



### 6. Output data

```{r}
tidy_data %>%
    write_csv("additional_files/2020_W2_Output_From_R")
```




  

