---
title: "2022 Week 3 preppin' data Challenge - R Solution"
author: "Paula Munoz"
date: "2022/01/26"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2023: Week 3 -The Prep School - Passing Grades

Challenge by Carl Allchin.

Challenge Link:
<https://preppindata.blogspot.com/2022/01/2022-week-3-prep-school-passing-grades.html>

Data Sources:
<https://drive.google.com/file/d/1p8gt3cR3ATCeGK81pnT90x0a6dbCXst1/view>
<https://drive.google.com/file/d/135o1Kj9koWM5eZ4VJjH9EUCyo1UnP54C/view>

### DETAILS

#### Requirements

- Input both data sets
- Join the data sets together to give us the grades per student
- Remove the parental data fields, they aren't needed for the challenge this week
- Pivot the data to create one row of data per student and subject
- Rename the pivoted fields to Subject and Score 
- Create an average score per student based on all of their grades
- Create a field that records whether the student passed each subject: Pass mark is 75 and above in all subjects
- Aggregate the data per student to count how many subjects each student passed
- Round the average score per student to one decimal place
- Remove any unnecessary fields and output the data


## LOADING LIBRARIES 

```{r pressure, message=FALSE}
library(tidyverse)
library(skimr)
library(knitr)
library(tidyquant)
library(lubridate)
```


## IMPORTING DATA 

Data resides in a .csv file

```{r}
input <- readr::read_csv('additional_files/PD 2022 Wk 1 Input - Input.csv', 
                           col_types = 
                             cols(
                                id = col_character()
                              ))
input_2 <- readr::read_csv('additional_files/PD 2022 WK 3 Grades.csv', 
                           col_types = 
                             cols(
                               `Student ID` = col_character()
                             ))

```


## INSPECTING DATA 

* Inspecting first five rows:
```{r}
input %>%
  head() %>%
  kable()

input_2 %>%
  head() %>%
  kable()
```

* Glimpse both tables:
```{r}
input %>%
  glimpse()

input_2 %>%
  glimpse()
```


## DATA WRANGLING

### Remove unnecessary columns
```{r}
# Keeping only Student ID and Gender from first table
tidy_data  <- input %>% select(id,gender)

tidy_data%>%
  head(10) %>%
  kable() 
```



### Join the datasets

```{r}

input_2 <- input_2 %>%
  #Renaming Student ID column under second dataset to id
  rename(id = `Student ID`)

tidy_data <- tidy_data %>%
  left_join(input_2)


tidy_data%>%
  head(10) %>%
  kable() 
```

### Pivot the data to create one row of data per student and subject
### Rename the pivoted fields to Subject and Score 

```{r}
tidy_data <- tidy_data %>%
  pivot_longer(Maths:Geography,
               names_to = "Subject",
               values_to = "Score")
```


### Create a field that records whether the student passed each subject: 
### Pass mark is 75 and above in all subjects

```{r}

tidy_data <- tidy_data %>%
  mutate(Subject_Passed = case_when(Score >= 75 ~ 1,
                                    TRUE ~ 0))
  
tidy_data%>%
  head(10) %>%
  kable() 
```

### Create an average score per student based on all of their grades
### Aggregate the data per student to count how many subjects each student passed
### Round the average score per student to one decimal place


```{r}

tidy_data <- tidy_data %>%
  group_by(id,gender) %>%
  summarise(Suject_Passed= sum(Subject_Passed),
            Students_Avg_Score = mean(Score)) %>%
  ungroup()
 
           
tidy_data%>%
  head(10) %>%
  kable()            

```

### Remove any unnecessary fields and output the data

```{r}
tidy_data <- tidy_data %>%
  mutate(Students_Avg_Score = round(Students_Avg_Score,1))

tidy_data%>%
  head(10) %>%
  kable() 
```


### Outut data

```{r}
tidy_data %>%
    write_csv("additional_files/2020_W3_Output_From_R")
```




  

