---
title: "2022 Week 14 preppin' data Challenge - R Solution"
author: "Paula Munoz"
date: "2022/04/10"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2022: Week 14 -House of Games Winners


Challenge by Jenny Martin.

Challenge Link:
<https://preppindata.blogspot.com/2022/04/2022-week-14-house-of-games-winners.html>

Data Sources:
<https://drive.google.com/file/d/1cKZbluTPNlbBHm5LWHDEVVdsD9SA8l-A/view>


### DETAILS

#### Requirements

- Input the data

- Only keep relevant fields and rename certain fields to remove duplication

-> Ser. becomes Series

-> Wk. becomes Week

-> T becomes Tu

-> T 1 becomes Th

-> Total becomes Score

-> Week becomes Points

-> Week 1 becomes Rank

- Filter the data to remove Series that have a null value, or are preceded by an 'N'

- Calculate the Points without double points Friday

-> Rank the players based on this new field

-> Create a field to determine if there has been a change in winner for that particular Series and Week

- Rank the players based on their Score instead

-> Create a field to determine if there has been a change in winner for that particular Series and Week

- Calculate the Score if the score on Friday was doubled (instead of the Points)

-> Rank the players based on this new field

-> Create a field to determine if there has been a change in winner for that particular Series and Week

- Remove unnecessary fields

- Output the data


## LOADING LIBRARIES 

```{r pressure, message=FALSE}
library(tidyverse)
library(skimr)
library(knitr)
library(tidyquant)
library(lubridate)
```


## IMPORTING DATA 

Data resides in a .csv file

```{r}
input <- readr::read_csv("additional_files/Richard Osman's House of Games - Episode Guide - Players.csv")
```


## INSPECTING DATA 

* Inspecting first five rows:
```{r}
input %>%
  head() %>%
  kable()

```

* Glimpse both tables:
```{r}
input %>%
  glimpse()

```

```{r}
dim(input)
```


## DATA WRANGLING

### Keep only relevant fields
```{r}
tidy_data  <- input %>%
  select(Player, Ser., Wk., Week, Total, Week_1, F, M_1, T_2, W_1, T_3,F_1)

tidy_data%>%
  head(10) %>%
  kable() 
```



### Rename fields
```{r}

tidy_data  <- tidy_data%>%
  rename(
    Rank    = Week_1,
    Points  = Week,
    Score   = Total,
    Week    = Wk.,
    Series  = Ser.
    )

tidy_data%>%
  head(10) %>%
  kable() 
```



### Filter the data to remove Series that have a null value, or are preceded by an 'N'

```{r}
# See different values for 'Series', we can see there is 'N1', 'N2' and Nulls/ NA
tidy_data %>%
  group_by(Series) %>%
  summarise( n = n())


```
```{r}
tidy_data <- tidy_data %>%
  filter(
    !is.na(Series),
    !str_detect(Series, "^N"))
```

```{r}
# Confirming unwanted values have been removed
tidy_data %>%
  group_by(Series) %>%
  summarise( n = n())


```
```{r}
dim(tidy_data)
# Reduced number of rows from 303 to 284
```

```{r}
tidy_data %>%
  glimpse()
```

### Convert Ranks to numeric


```{r}
 tidy_data$Rank <- tidy_data$Rank %>%
  substr(1,1) %>%
  as.integer()
  
 tidy_data$M_1 <- tidy_data$M_1 %>%
  substr(1,1) %>%
  as.integer()
 
  tidy_data$T_2 <- tidy_data$T_2 %>%
  substr(1,1) %>%
  as.integer()
  
  tidy_data$W_1 <- tidy_data$W_1 %>%
  substr(1,1) %>%
  as.integer()

 tidy_data$T_3 <- tidy_data$T_3 %>%
  substr(1,1) %>%
  as.integer()
  
 tidy_data$F_1 <- tidy_data$F_1 %>%
  substr(1,1) %>%
  as.integer()

tidy_data%>%
  glimpse()  
```


```{r}
tidy_data %>%
  head()
```
### Calculate the Points without double points Friday
```{r}

tidy_data <- tidy_data %>%
  mutate(points_without_double_points_friday = (5 - M_1) + (5 - T_2) + (5 - W_1) + (5 - T_3) + (5 - F_1))

tidy_data%>%
  head(10) %>%
  kable() 
```
### Rank without double points Friday
```{r}
tidy_data <- tidy_data %>%
  group_by(Series, Week) %>%
  mutate(rank_without_double_points_friday = dense_rank(desc(points_without_double_points_friday))) %>%
  ungroup()

```

## Create a field to determine if there has been a change in winner for that particular Series and Week

```{r}
tidy_data_change_1 <- tidy_data %>%
  group_by(Series, Week) %>%
  arrange(Rank) %>%
  mutate(change_in_winner_no_double_points_friday =  case_when(Rank == rank_without_double_points_friday ~ 'FALSE',
                                                               TRUE ~ 'TRUE')) %>%
  filter(row_number() == 1) %>%
  select(Series, Week, change_in_winner_no_double_points_friday) 
```

## Joining to original table and including change_in_winner_no_double_points_friday
```{r}
tidy_data <- tidy_data %>%
  left_join(tidy_data_change_1)

tidy_data%>%
  head(10) %>%
  kable()
```
### Rank the players based on their Score instead

```{r}
tidy_data <- tidy_data %>%
  group_by(Series, Week) %>%
  mutate(rank_based_on_score = dense_rank(desc(Score))) %>%
  ungroup()
```


### Create a field to determine if there has been a change in winner for that particular Series and Week
```{r}
tidy_data_change_2 <- tidy_data %>%
  group_by(Series, Week) %>%
  arrange(Rank) %>%
  mutate(change_in_winner_based_on_score =  case_when(Rank == rank_based_on_score ~ 'FALSE',
                                                               TRUE ~ 'TRUE')) %>%
  filter(row_number() == 1) %>%
  select(Series, Week, change_in_winner_based_on_score) 
```

## Joining to original table and including change_in_winner_based_on_score
```{r}
tidy_data <- tidy_data %>%
  left_join(tidy_data_change_2)

tidy_data%>%
  head(10) %>%
  kable()
```

### Calculate the Score if the score on Friday was doubled (instead of the Points)

```{r}
tidy_data <- tidy_data %>%
  mutate(score_if_double_friday = Score + F)

tidy_data%>%
  head(10) %>%
  kable()
```

### Rank the players based on this new field
```{r}
tidy_data <- tidy_data %>%
  group_by(Series, Week) %>%
  mutate(rank_if_double_score_friday = dense_rank(desc(score_if_double_friday))) %>%
  ungroup()
```

### Create a field to determine if there has been a change in winner for that particular Series and Week
```{r}
tidy_data_change_3 <- tidy_data %>%
  group_by(Series, Week) %>%
  arrange(Rank) %>%
  mutate(change_in_winner_double_score_friday=  case_when(Rank == rank_if_double_score_friday ~ 'FALSE',
                                                               TRUE ~ 'TRUE')) %>%
  filter(row_number() == 1) %>%
  select(Series, Week, change_in_winner_double_score_friday) 
```

## Joining to original table and including change_in_winner_double_score_friday
```{r}
tidy_data <- tidy_data %>%
  left_join(tidy_data_change_3)

tidy_data%>%
  head(10) %>%
  kable()
```


### Remove unnecessary fields

```{r}

tidy_data <- tidy_data %>%
  select(Series, Week, Player, Rank, rank_without_double_points_friday, change_in_winner_no_double_points_friday, rank_based_on_score, change_in_winner_based_on_score, rank_if_double_score_friday, change_in_winner_double_score_friday, Score, Points, score_if_double_friday, points_without_double_points_friday) %>%
  arrange(Series, Week)
  
 tidy_data%>%
  head(10) %>%
  kable()  
```




### Outut data

```{r}
tidy_data %>%
    write_csv("additional_files/2022_W14_Output_From_R.csv")
```




  

