---
title: "2021 W 32 TidyTuesday - Paralympics"
author: "Paula Munoz"
date: "08/08/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2021: Week 32 - Paralympics

Challenge by @R4DScommunity 

Challenge Link: <https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-08-03>.

Data Source: <https://db.ipc-services.org/sdms/hira/web/index>

### DETAILS

Data webscraped for the games of 1980 - 2016 for the athletes for all of their listed sports. There are additional summary-level datasets for medals by country/sport etc that are also available on the site.



## LOADING LIBRARIES 

```{r pressure, message=FALSE}
library(tidyverse)
library(skimr)
library(knitr)
library(tidyquant)
library(rvest)
library(ggthemes)

```


## IMPORTING DATA 


Importing data from Github Repository

```{r}
athletes <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-03/athletes.csv')


```


## INSPECTING DATA 

* Inspecting first five rows:
```{r}
athletes %>%
  head() %>%
  kable()
```



* Detailed summary:
```{r}
athletes %>%
  skim()
```


* Glimpse data:
```{r}
athletes %>%
  glimpse()
```

```{r}
athletes  %>%
  count(country)
```
### CLEANING SCRIPT - Provided by @R4DScommunity 

To see full cleaning code go to TidyTuesday page, I'm excluding it from here for simplicity purposes.

```{r,echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
core_url <- "https://db.ipc-services.org/sdms/hira/web/competition/code/PG2016/sport/AR"

"body > div.container.bg-white.border.border-header.border-top-0.pb-3 > table > tbody > tr:nth-child(1) > td:nth-child(7) > div > a"

raw_html <- "https://db.ipc-services.org/sdms/hira/web/competition/code/PG2016" %>% 
  read_html()


# Archery -----------------------------------------------------------------


clean_arrow <- function(year){
  
  arrows <- glue::glue("https://db.ipc-services.org/sdms/hira/web/competition/code/PG{year}/sport/AR") %>% 
    read_html()
  
  raw_arrow <- arrows %>% 
    html_table() %>% 
    .[[2]] %>% 
    pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>% 
    rename(event = Event) %>% 
    mutate(
      medal = str_remove(medal, " Medallist\\(s\\)")
    )
  
  ind_arrow <- raw_arrow %>% 
    separate(athlete, into = c("athlete", "abb"), sep = " \\(") %>% 
    separate(event, into = c("gender", "event"), sep = "'s | ", fill = "left", extra = "merge") %>% 
    mutate(abb = str_remove(abb, "\\)")) %>% 
    filter(str_detect(event, "Mixed|Team", negate = TRUE)) 
  
  team_arrow <- raw_arrow %>% 
    filter(str_detect(event, "Mixed|Team")) %>% 
    separate(athlete, into = c("country", "abb"), sep = " \\(") %>% 
    separate(abb, into = c("abb", "athlete"), sep = "\\)") %>% 
    separate(event, into = c("gender", "event"), sep = " |'s ", fill = "left", extra = "merge") %>%
    separate_rows(athlete, sep = "(?<=[a-z])(?=[A-Z])") %>% 
    group_by(event, medal, abb) %>% 
    mutate(grp_id = row_number()) %>% 
    ungroup()
  
  bind_rows(ind_arrow, team_arrow) %>% 
    mutate(type = "Archery", year = year) 
}

clean_arrow(2008) %>% 
  print(n = 100)

year_vec <- seq(1980, 2016, by = 4)

try_arrow <- safely(clean_arrow)

all_arrow <- year_vec %>% 
  map(try_arrow) 

arrow_years <- all_arrow %>% 
  map_dfr("result")

# athletics ---------------------------------------------------------------

sport_vec <- raw_html %>% 
  html_nodes("td:nth-child(7) > div > a") %>% 
  html_attr("href") %>% 
  str_remove("/sdms/hira/web/competition/code/PG2016/sport/")


clean_ath <- function(year){
  
  ath <- glue::glue("https://db.ipc-services.org/sdms/hira/web/competition/code/PG{year}/sport/AT") %>% 
    read_html()
  
  raw_ath <- ath %>% 
    html_table() %>% 
    .[[2]] %>% 
    pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>% 
    rename(event = Event) %>% 
    mutate(
      medal = str_remove(medal, " Medallist\\(s\\)")
    ) %>% 
    separate(event, into = c("gender", "event"), sep = " |'s ", fill = "left", extra = "merge")
  
  raw_ind <- raw_ath %>% 
    filter(str_detect(event, "4x", negate = TRUE)) %>% 
    separate(athlete, into = c("athlete", "abb"), sep = " \\(", fill = "left", extra = "merge") %>% 
    separate(abb, into = c("abb", "guide"), sep = "\\)") %>% 
    mutate(guide = str_remove(guide, "Guide: "),
           guide = if_else(guide == "", NA_character_, guide)) %>% 
    filter(!is.na(athlete)) 
  
  clean_ind <- raw_ind %>% 
    add_row(
      raw_ind %>% 
        filter(str_detect(guide, "\\(")) %>% 
        select(gender:medal, athlete = guide) %>% 
        separate(athlete, c("athlete", "abb"), sep = " \\(")
    ) %>% 
    filter(!is.na(athlete)) %>% 
    mutate(guide = if_else(str_detect(guide, "\\("), NA_character_, guide)) 
  
  
  clean_grp <- raw_ath %>% 
    filter(str_detect(event, "4x")) %>% 
    separate(athlete, into = c("country", "abb", "athlete"), sep = " \\(|\\)") %>% 
    separate_rows(athlete, sep = "(?<=[a-z])(?=[A-Z])") %>% 
    separate_rows(athlete, sep = "(?<=[A-Z]\\.)(?=[A-Z])") %>% 
    mutate(
      guide = if_else(
        str_detect(lead(athlete), "Guide"), 
        lead(athlete),
        NA_character_
      )
    ) %>% 
    filter(str_detect(athlete, "Guide", negate = TRUE)) %>%
    group_by(event, medal, country) %>% 
    mutate(grp_id = row_number()) %>% 
    ungroup()
  
  bind_rows(clean_ind, clean_grp) %>% 
    mutate(type = "Athletics", year = year)
  
}

test_df <- clean_ath(1980) 
test_df %>% 
  filter(str_detect(event, "4x60")) %>% 
  print(n = 100)

safe_ath <- safely(clean_ath)

ath_years <- year_vec %>% 
  map(safe_ath) %>% 
  map_dfr("result")

# Cycling -----------------------------------------------------------------

clean_cycle <- function(year){
  
  url_c <- glue::glue("https://db.ipc-services.org/sdms/hira/web/competition/code/PG{year}/sport/CY") %>% 
    read_html()
  
  raw_cycle <- url_c %>% 
    html_table() %>% 
    .[[2]] %>% 
    pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>% 
    rename(event = Event) %>% 
    mutate(
      medal = str_remove(medal, " Medallist\\(s\\)")
    ) %>% 
    separate(event, into = c("gender", "event"), sep = " |'s ", fill = "left", extra = "merge")
  
  raw_ind_c <- raw_cycle %>% 
    filter(str_detect(event, "Team|Tandem", negate = TRUE)) %>% 
    separate(athlete, into = c("athlete", "abb"), sep = " \\(", fill = "left", extra = "merge") %>% 
    separate(abb, into = c("abb", "pilot"), sep = "Pilot: |\\(PT\\):") %>% 
    separate(athlete, sep = "(?<=[PT\\)])(?=[A-Z])", into = c("athlete", "pilot")) %>%
    separate(athlete, sep = "(?<=[a-z]+)(?=[A-Z])", into = c("athlete", "pilot")) %>%
    mutate(pilot = if_else(pilot == "", NA_character_, pilot)) %>% 
    filter(!is.na(athlete)) 
  
  clean_ind_c <- raw_ind_c %>% 
    add_row(
      raw_ind_c %>% 
        filter(str_detect(pilot, "\\(")) %>% 
        select(gender:medal, athlete = pilot) %>% 
        separate(athlete, c("athlete", "abb"), sep = " \\(")
    ) %>% 
    filter(!is.na(athlete)) %>% 
    mutate(pilot = if_else(str_detect(pilot, "\\("), NA_character_, pilot)) %>% 
    mutate(abb = str_remove(abb, "\\)"))
  
  
  clean_grp_c <- raw_cycle %>% 
    filter(str_detect(event, "Team")) %>% 
    separate(athlete, into = c("country", "abb", "athlete"), sep = " \\(|\\)") %>% 
    separate_rows(athlete, sep = "(?<=[a-z])(?=[A-Z])") %>% 
    mutate(
      guide = if_else(
        str_detect(lead(athlete), "Pilot|PT\\)"), 
        lead(athlete),
        NA_character_
      )
    ) %>% 
    filter(str_detect(athlete, "Pilot|PT\\)", negate = TRUE)) %>%
    group_by(event, medal, country) %>% 
    mutate(grp_id = row_number()) %>% 
    ungroup()
  
  bind_rows(clean_ind_c, clean_grp_c) %>% 
    mutate(year = year, type = "Cycling") %>% 
    rename(guide = pilot)
  
}

safe_cycle <- safely(clean_cycle)

cycle_years <- year_vec %>% 
  map(safe_cycle) %>% 
  map_dfr("result")


# Powerlifting ------------------------------------------------------------


clean_power <- function(year){
  
  raw_power <- glue::glue("https://db.ipc-services.org/sdms/hira/web/competition/code/PG{year}/sport/PO") %>% 
    read_html()
  
  clean_pwr <- raw_power %>% 
    html_table() %>% 
    .[[2]] %>% 
    pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>% 
    rename(event = Event) %>% 
    mutate(
      medal = str_remove(medal, " Medallist\\(s\\)")
    ) %>% 
    separate(event, into = c("gender", "event"), sep = " |'s ", fill = "left", extra = "merge") %>% 
    separate(athlete, into = c("athlete", "abb"), sep = " \\(", fill = "left", extra = "merge") %>% 
    separate_rows(abb, sep = " \\)") 
  
  clean_abb_power <- clean_pwr %>%
    mutate(abb = str_remove(abb, "\\)")) %>%
    bind_rows(clean_pwr %>%
                filter(str_detect(abb, "\\(")) %>%
                separate_rows(abb, sep = "\\(") %>%
                mutate(athlete = ifelse(str_length(abb) <= 4, lag(abb), athlete)) %>%
                mutate(
                  abb = str_remove(abb, "\\).*"),
                  athlete = str_remove(athlete, ">*\\)") %>% str_trim()
                )) %>%
    filter(str_detect(abb, "\\(", negate = TRUE), !is.na(athlete)) %>% 
    mutate(abb = str_remove(abb, "\\)"))
  
  
  clean_abb_power %>% 
    mutate(year = year, type = "Powerlifting")
}

clean_power(1984)

safe_power <- safely(clean_power)

power_years <- year_vec %>% 
  map(safe_power) %>% 
  map_dfr("result")


# Swimming ----------------------------------------------------------------

clean_swim <- function(year){
  
  raw_swim <- glue::glue("https://db.ipc-services.org/sdms/hira/web/competition/code/PG{year}/sport/SW") %>%
  read_html()

raw_ind_sw <- raw_swim %>%
  html_table() %>%
  .[[2]] %>%
  pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>%
  rename(event = Event) %>%
  mutate(
    medal = str_remove(medal, " Medallist\\(s\\)")
  ) %>%
  filter(str_detect(event, "4x", negate = TRUE)) %>%
  separate(event, into = c("gender", "event"), sep = " |'s ", fill = "left", extra = "merge") %>% 
  separate_rows(athlete, sep = "(?<=[A-Z]\\))(?=[A-Z])") %>%
  separate(athlete, into = c("athlete", "abb"), sep = " \\(", fill = "left", extra = "merge")


clean_ind_sw <- raw_ind_sw %>%
  mutate(abb = str_remove(abb, "\\)")) %>%
  bind_rows(raw_ind_sw %>%
    filter(str_detect(abb, "\\(")) %>%
    separate_rows(abb, sep = "\\(") %>%
    mutate(athlete = ifelse(str_length(abb) <= 4, lag(abb), athlete)) %>%
    mutate(
      abb = str_remove(abb, "\\).*"),
      athlete = str_remove(athlete, ">*\\)") %>% str_trim()
    )) %>%
    filter(str_detect(abb, "\\(", negate = TRUE), !is.na(athlete))


clean_grp_sw <- raw_swim %>%
  html_table() %>%
  .[[2]] %>%
  pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>%
  rename(event = Event) %>%
  mutate(
    medal = str_remove(medal, " Medallist\\(s\\)")
  ) %>%
  filter(str_detect(event, "4x")) %>%
  separate(event, into = c("gender", "event"), sep = " |'s ", fill = "left", extra = "merge") %>%
  separate(athlete, into = c("country", "abb", "athlete"), sep = " \\(|\\)") %>%
  separate_rows(athlete, sep = "(?<=[a-z])(?=[A-Z])") %>%
  group_by(event, medal, country) %>%
  mutate(grp_id = row_number()) %>%
  ungroup()

bind_rows(clean_ind_sw, clean_grp_sw) %>% 
  mutate(year = year, type = "Swimming") 

  }

safe_swim <- safely(clean_swim)

swim_years <- year_vec %>% 
  map(safe_swim) %>% 
  map_dfr("result")



# Table Tennis ------------------------------------------------------------

clean_tab_tennis <- function(year){
  
  raw_tab <- glue::glue("https://db.ipc-services.org/sdms/hira/web/competition/code/PG{year}/sport/TT") %>%
    read_html()
  
  clean_ind_tab <- raw_tab %>%
    html_table() %>%
    .[[2]] %>%
    pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>%
    rename(event = Event) %>%
    mutate(
      medal = str_remove(medal, " Medallist\\(s\\)")
    ) %>%
    filter(str_detect(event, "Team", negate = TRUE)) %>%
    separate(
      event,
      into = c("gender", "event"),
      sep = " |'s ",
      fill = "left",
      extra = "merge"
    ) %>%
    separate_rows(athlete, sep = "(?<=[\\)])(?=[A-Z])") %>%
    separate(
      athlete,
      into = c("athlete", "abb"),
      sep = " \\(",
      fill = "left",
      extra = "merge"
    ) %>%
    mutate(abb = str_remove(abb, "\\)"))
  
  
  clean_grp_tab <- raw_tab %>%
    html_table() %>%
    .[[2]] %>%
    pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>%
    rename(event = Event) %>%
    mutate(
      medal = str_remove(medal, " Medallist\\(s\\)")
    ) %>%
    filter(str_detect(event, "Team")) %>%
    separate(event, into = c("gender", "event"), sep = " |'s ", fill = "left", extra = "merge") %>%
    separate(athlete, into = c("country", "abb", "athlete"), sep = " \\(|\\)") %>%
    separate_rows(athlete, sep = "(?<=[a-z])(?=[A-Z])") %>%
    separate_rows(athlete, sep = "(?<=[A-Z]\\.)(?=[A-Z])") %>% 
    group_by(event, medal, country) %>%
    mutate(grp_id = row_number()) %>%
    ungroup()
  
  bind_rows(clean_ind_tab, clean_grp_tab) %>% 
    mutate(year = year, type = "Table Tennis") 
  
}

safe_table <- safely(clean_tab_tennis)

table_years <- year_vec %>% 
  map(safe_table) %>% 
  map_dfr("result")



# Volleyball --------------------------------------------------------------



clean_volleyball <- function(year){
  
  raw_vb <- glue::glue("https://db.ipc-services.org/sdms/hira/web/competition/code/PG{year}/sport/VO") %>%
    read_html()
  
  clean_grp_vb <- raw_vb %>%
    html_table() %>%
    .[[2]] %>%
    pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>%
    rename(event = Event) %>%
    mutate(
      medal = str_remove(medal, " Medallist\\(s\\)")
    ) %>%
    separate(event, into = c("gender", "event"), sep = " |'s ", fill = "left", extra = "merge") %>%
    separate(athlete, into = c("country", "abb", "athlete"), sep = " \\(|\\)") %>%
    separate_rows(athlete, sep = "(?<=[a-z])(?=[A-Z])") %>%
    separate_rows(athlete, sep = "(?<=[A-Z]\\.)(?=[A-Z])") %>% 
    group_by(event, medal, country) %>%
    mutate(grp_id = row_number()) %>%
    ungroup()
  
  clean_grp_vb %>% 
    mutate(year = year, type = "Volleyball")
  
}

safe_vb <- safely(clean_volleyball)

vb_years <- year_vec %>% 
  map(safe_vb) %>% 
  map_dfr("result")



# Basketball --------------------------------------------------------------

clean_basketball <- function(year){
  
  raw_bb <- glue::glue("https://db.ipc-services.org/sdms/hira/web/competition/code/PG{year}/sport/WB") %>%
    read_html()
  
  clean_grp_bb <- raw_bb %>%
    html_table() %>%
    .[[2]] %>%
    pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>%
    rename(event = Event) %>%
    mutate(
      medal = str_remove(medal, " Medallist\\(s\\)")
    ) %>%
    separate(event, into = c("gender", "event"), sep = " |'s ", fill = "left", extra = "merge") %>%
    separate(athlete, into = c("country", "abb", "athlete"), sep = " \\(|\\)") %>%
    separate_rows(athlete, sep = "(?<=[a-z])(?=[A-Z])") %>%
    group_by(event, medal, country) %>%
    mutate(grp_id = row_number()) %>%
    ungroup()
  
  clean_grp_bb %>% 
    mutate(year = year, type = "Basketball")
  
}

safe_bb <- safely(clean_basketball)

bb_years <- year_vec %>% 
  map(safe_bb) %>% 
  map_dfr("result")



# Fencing -----------------------------------------------------------------

clean_fencing <- function(year){
  
  raw_f <- glue::glue("https://db.ipc-services.org/sdms/hira/web/competition/code/PG{year}/sport/WF") %>%
    read_html()
  
  clean_ind_f <- raw_f %>%
    html_table() %>%
    .[[2]] %>%
    pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>%
    rename(event = Event) %>%
    mutate(
      medal = str_remove(medal, " Medallist\\(s\\)")
    ) %>%
    filter(str_detect(event, "Team", negate = TRUE)) %>%
    separate(event, into = c("gender", "event"), sep = " |'s ", fill = "left", extra = "merge") %>%
    separate(athlete, into = c("athlete", "abb"), sep = " \\(", fill = "left", extra = "merge") %>% 
    mutate(abb = str_remove(abb, "\\)"))
  
  clean_grp_f <- raw_f %>%
    html_table() %>%
    .[[2]] %>%
    pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>%
    rename(event = Event) %>%
    mutate(
      medal = str_remove(medal, " Medallist\\(s\\)")
    ) %>%
    filter(str_detect(event, "Team")) %>%
    separate(event, into = c("gender", "event"), sep = " |'s ", fill = "left", extra = "merge") %>%
    separate(athlete, into = c("country", "abb", "athlete"), sep = " \\(|\\)") %>%
    separate_rows(athlete, sep = "(?<=[a-z])(?=[A-Z])") %>%
    separate_rows(athlete, sep = "(?<=[A-Z]\\.)(?=[A-Z])") %>% 
    group_by(event, medal, country) %>%
    mutate(grp_id = row_number()) %>%
    ungroup()
  
  bind_rows(clean_ind_f, clean_grp_f) %>% 
    mutate(year = year, type = "Fencing")
  
}
  
safe_fence <- safely(clean_fencing)

fence_years <- year_vec %>% 
  map(safe_fence) %>% 
  map_dfr("result")




# Rugby -------------------------------------------------------------------

clean_rugby <- function(year){
  
  raw_rug <- glue::glue("https://db.ipc-services.org/sdms/hira/web/competition/code/PG{year}/sport/WR") %>%
    read_html()
  
  clean_grp_rug <- raw_rug %>%
    html_table() %>%
    .[[2]] %>%
    pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>%
    rename(event = Event) %>%
    mutate(
      medal = str_remove(medal, " Medallist\\(s\\)"),
      event = paste(event, "Wheelchair Rugby")
    ) %>%
    separate(event, into = c("gender", "event"), sep = " |'s ", fill = "left", extra = "merge") %>%
    separate(athlete, into = c("country", "abb", "athlete"), sep = " \\(|\\)") %>%
    separate_rows(athlete, sep = "(?<=[a-z])(?=[A-Z])") %>%
    separate_rows(athlete, sep = "(?<=[A-Z]\\.)(?=[A-Z])") %>% 
    group_by(event, medal, country) %>%
    mutate(grp_id = row_number()) %>% 
    ungroup()
  
  clean_grp_rug %>% 
    mutate(year = year, type = "Rugby")
  
}

safe_rugby <- safely(clean_rugby)

rugby_years <- year_vec %>% 
  map(safe_rugby) %>% 
  map_dfr("result")




# Tennis ------------------------------------------------------------------

clean_tennis <- function(year){
  
  raw_ten <- glue::glue("https://db.ipc-services.org/sdms/hira/web/competition/code/PG{year}/sport/TT") %>%
    read_html()
  
  clean_ind_ten <- raw_ten %>%
    html_table() %>%
    .[[2]] %>%
    pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>%
    rename(event = Event) %>%
    mutate(
      medal = str_remove(medal, " Medallist\\(s\\)")
    ) %>%
    filter(str_detect(event, "Double|Team", negate = TRUE)) %>%
    separate(event, into = c("gender", "event"), sep = " |'s ", fill = "left", extra = "merge") %>%
    separate_rows(athlete, sep = "(?<=[\\)])(?=[A-Z])") %>%
    separate(athlete, into = c("athlete", "abb"), sep = " \\(", fill = "left", extra = "merge") %>% 
    mutate(abb = str_remove(abb, "\\)"))
  
  
  clean_grp_ten <- raw_ten %>%
    html_table() %>%
    .[[2]] %>%
    pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>%
    rename(event = Event) %>%
    mutate(
      medal = str_remove(medal, " Medallist\\(s\\)")
    ) %>%
    filter(str_detect(event, "Double|Team")) %>%
    separate(event, into = c("gender", "event"), sep = " |'s ", fill = "left", extra = "merge") %>%
    separate(athlete, into = c("country", "abb", "athlete"), sep = " \\(|\\)") %>%
    separate_rows(athlete, sep = "(?<=[a-z])(?=[A-Z])") %>%
    group_by(event, medal, country) %>%
    mutate(grp_id = row_number()) %>%
    ungroup()
  
  bind_rows(clean_ind_ten, clean_grp_ten) %>% 
    mutate(year = year, type = "Wheelchair Tennis")
  
}

safe_tennis <- safely(clean_tennis)

tennis_years <- year_vec %>% 
  map(safe_tennis) %>% 
  map_dfr("result")



# Triathlon ---------------------------------------------------------------

clean_triathlon <- function(year){
  
  tri_url <- glue::glue("https://db.ipc-services.org/sdms/hira/web/competition/code/PG{year}/sport/TR") %>% 
    read_html()
  
  raw_tri <- tri_url %>% 
    html_table() %>% 
    .[[2]] %>% 
    pivot_longer(names_to = "medal", values_to = "athlete", cols = -Event) %>% 
    rename(event = Event) %>% 
    mutate(
      medal = str_remove(medal, " Medallist\\(s\\)")
    ) %>% 
    separate(event, into = c("gender", "event"), sep = " |'s ", fill = "left", extra = "merge")
  
  raw_ind_tri <- raw_tri %>% 
    filter(str_detect(event, "4x", negate = TRUE)) %>% 
    separate(athlete, into = c("athlete", "abb"), sep = " \\(", fill = "left", extra = "merge") %>% 
    separate(abb, into = c("abb", "guide"), sep = "\\)") %>% 
    mutate(guide = str_remove(guide, "Guide: "),
           guide = if_else(guide == "", NA_character_, guide)) %>% 
    filter(!is.na(athlete)) 
  
  clean_ind_tri <- raw_ind_tri %>% 
    add_row(
      raw_ind_tri %>% 
        filter(str_detect(guide, "\\(")) %>% 
        select(gender:medal, athlete = guide) %>% 
        separate(athlete, c("athlete", "abb"), sep = " \\(")
    ) %>% 
    filter(!is.na(athlete)) %>% 
    mutate(guide = if_else(str_detect(guide, "\\("), NA_character_, guide)) %>% 
    mutate(year = year, type = "Triathlon")
  
  
  clean_ind_tri
}

safe_tri <- safely(clean_triathlon)

tri_years <- year_vec %>% 
  map(safe_tri) %>% 
  map_dfr("result")



all_sports <- bind_rows(
  list(arrow_years, ath_years, bb_years, cycle_years, fence_years, power_years,
       rugby_years, swim_years, table_years, tennis_years, tri_years, vb_years)
)

all_sports %>% 
  skimr::skim()

all_sports %>% 
  write_csv("athletes2.csv")



check_abb <- function(df_in) {
df_in %>% 
    filter(str_length(abb) > 3) %>% 
    select(athlete, abb, type, year)
  
}

all_sports %>% check_abb() %>% distinct(type, year)


```


```{r}
all_sports %>%
  glimpse()
```


## DATA WRANGLING

### 1. Selecting columns of interest and summarizing

```{r}
tidy_data <- all_sports %>%
  group_by(abb, medal) %>%
  summarise(count_medals = n()) %>%
  filter(abb != "-") %>%
  ungroup()



tidy_data%>%
  head(10) %>%
  kable()
```

### 2. Importing dataset breaking down Country by Continent

I would like to visulize medals won by continent, since continent is not part of original dataset, I searched for a dataset to complement my data.

Source of data including Country and Continent list: <https://datahub.io/JohnSnowLabs/country-and-continent-codes-list>
```{r}
continent_Country <- read_csv("continent_country.csv") %>%
   select(Continent_Name, Country_Name, Three_Letter_Country_Code)

continent_Country %>%
  head() %>%
  kable()

```

### 3. Join sports tidy data with continent_Country
```{r}

tidy_data_continent <- tidy_data %>%
  left_join(continent_Country, by = c("abb" = "Three_Letter_Country_Code") ) %>%
  filter(Continent_Name != " ")

tidy_data_continent  %>%
  head() %>%
  kable()
```

### 2. Create stacked bar
```{r, fig.asp = .62}
level_order <- factor(tidy_data_continent$medal, level = c("Silver", "Bronze", "Gold"))

ggplot(tidy_data_continent, aes(fill=level_order, y=count_medals, x=Continent_Name)) +
  geom_bar(position="fill", stat = "identity") + coord_flip() +
  scale_fill_manual(values = c(
    "Bronze" = "#A77044",
    "Gold" = "#D6AF36",
    "Silver" = "#A7A7AD"
  ))  + 
  labs(title = "Paralympics - Medals won by Continent 
       between 1980 - 2016", 
       subtitle = "Asia and North America are the continents that have 
       hitorically won more Gold Medals during the Paralympics",
    y = "Percent rate", 
    x = "Continent", 
    fill = "Type of Medal"
    )  +
  theme_fivethirtyeight() +
  theme(axis.title = element_text() )
  
  


```
